Sizing the bets in a focused portfolio

### 2023-05-17 doc/paper.tex
Address paper review comments

Before going on to explain the motivation about this work, a bit of context
might come useful. In silico is an employee--owned engineering consultancy
company that invests its excess cash into publicly traded stocks. The excess
cash comes sporadically, and In silico has been a net buyer over the last four
years: A trend that will likely continue. When the excess cash comes in, the
following question arises: How much money to put in which stocks? That question
had been answered by looking at our fundamental analysis for each company, and
performing some vague hand calculations. That of course works well because of
all the other uncertainties in the investment process, but the motivation was
to:
\begin{enumerate}
    \item Improve the odds of us behaving rationally (i.e. minimize
    psychological misjudgments such as anchoring bias~\cite{kahneman},
    consistency and commitment tendency~\cite{cialdini}, and
    others with their combined effects~\cite{almanack}),
    \item Save time (i.e. avoid doing hand calculations).
\end{enumerate}

@@

The answer to this question has been given by Kelly with his widely known Kelly
formula (sometimes called Kelly criterion)~\cite{kelly}. The Kelly's approach
starts by maximizing long--term growth of capital when one is presented with
infinite amount of opportunities to bet. This work extends that idea by
considering:
\begin{enumerate}
    \item Multiple companies (stocks) in parallel,
    \item An arbitrary number of scenarios for each company,
    \item A fundamentals--based analysis (after all, a stock is an ownership
    share of business~\cite{intelligentInvestor}.
\end{enumerate}

@@

During the mathematical derivation presented in \autoref{sec:mathematics}, an
assumption is made that the number of bets is very high (tends to infinity).
This work does not try to justify this assumption in a strong mathematical
form~\footnote{It is on the author's TODO list to try and prove this.}. Here's a
soft, non-mathematical reasoning on why your author thinks this is fine: The
assumption is made in order to write the equations in terms of probabilities
instead of the number of outcomes divided by the number of bets. Therefore, as
long as the input probabilities are \textit{conservatively} estimated, the
framework should still be valid. This essentially represents the most important
margin of safety~\cite{intelligentInvestor}.

@@

\noindent where $\mathcal{A}$ is the value of assets (capital), $N_c$ is the
number of candidate companies to consider, $f_j$ is the allocated fraction to
$j$th company, and $k_j$ is one of the returns for a company $j$ defined as the
relative difference between the estimated intrinsic value under a given scenario
($\mathcal{V}$) and the market capitalization at the time of investing
($\mathcal{M}$):
\begin{equation}
\label{eq:2}
    k_j = \frac{\mathcal{V}_j - \mathcal{M}_j}{\mathcal{M}_j}
\end{equation}

@@

\noindent where $n_i \in {n_{i1}, n_{i2}, \hdots, n_{N_o}} $ is the number of
the $i$th outcome that occurred $n_i$ times. Note that $k_{ij}$ represents the
return of the $j$th company in the $i$th outcome. Following original Kelly's
approach, a logarithmic growth function $\mathcal{G}$ is introduced:
\begin{equation}
\label{eq:4}
    \mathcal{G} = \lim_{N_a \to \infty} \frac{1}{N_a} \ln
        \frac{\mathcal{A}_{N_a}}{\mathcal{A}_0}
\end{equation}

@@

\noindent If one assumes an infinite number of (re)allocations $N_a$,
the following relation holds\footnote{This is an assumption that your author is
uncomfortable with, but \textit{feels} it is fine because of the margins of
safety embedded into the thinking that goes into assessing each investment
opportunity.}:
\begin{equation}
\label{eq:7}
    \lim_{N_a \to \infty} \frac{n_i}{N} = p_i
\end{equation}

@@

\begin{table}[!ht]
\caption{Company E with current market cap of 17.6 CAD.}
\vspace{0.25cm}
\centering
\begin{tabular}{l|c|c}
Scenario & Intrinsic value & Probability \\
\hline
Total loss & 0 CAD & 5\% \\
Bear thesis & 10M CAD & 25\% \\
Base thesis & 25M CAD & 70\% \\
\end{tabular}%
\label{tab:companyE}%
\end{table}%

\begin{table}[!ht]
\caption{Company F with current market cap of 581M USD.}
\vspace{0.25cm}
\centering
\begin{tabular}{l|c|c}
Scenario & Intrinsic value & Probability \\
\hline
Total loss & 0 USD & 20\% \\
Base thesis & 1B CAD & 80\% \\
\end{tabular}%
\label{tab:companyF}%
\end{table}%

@@

\noindent The last item is particularly interesting to your author. According to
Actuarial Life Tables in~\cite{lifeTables}, the probability of your currently 33
year--old author dying within the next year is approximately 0.25\%. That is two
orders of magnitude higher than the probability of the permanent loss of capital
for this portfolio. Considering that the portfolio has six stocks, that is a
very strong argument against excessive diversification, especially if:
\begin{itemize}
    \item One thinks of stocks as ownership shares of businesses, which implies
    long--term thinking and not being bothered by market fluctuations, 
    \item One embeds a margin of safety in different scenarios for different
    companies by e.g. recognizing that both unknown and known bad things may
    happen.
\end{itemize}

\noindent The observation about excessive diversification Poor Charlie's
Almanac~\cite{almanack}.

>>>>

Before going on to explain the motivation about this work, a bit of context
might come useful. In silico is an employee--owned engineering consultancy
company that invests its excess cash into publicly traded stocks. The excess
cash comes sporadically, and In silico has been a net buyer over the last four
years: A trend that will likely continue. When the excess cash comes in, the
following question arises: How much money to put in which stocks? That question
had been answered by looking at our fundamental analysis for each company, and
performing some vague hand calculations. That of course works well because of
all the other uncertainties in the investment process, but we were still
motivated to:
\begin{enumerate}
    \item Improve the odds of us behaving rationally (i.e. minimize
    psychological misjudgments such as anchoring bias~\cite{kahneman},
    consistency and commitment tendency~\cite{cialdini}, and
    others with their combined effects~\cite{almanack}),
    \item Save time (i.e. avoid doing hand calculations).
\end{enumerate}

@@

The answer to this question has been given by Kelly with his widely known Kelly
formula (sometimes called Kelly criterion)~\cite{kelly}. The Kelly's approach
starts by maximizing long--term growth of capital when one is presented with
an infinite amount of opportunities to bet on. This work extends that idea by
considering:
\begin{enumerate}
    \item Multiple companies (stocks) in parallel,
    \item An arbitrary number of scenarios for each company,
    \item A fundamentals--based analysis (after all, a stock is an ownership
    share of business~\cite{intelligentInvestor}).
\end{enumerate}

@@

During the mathematical derivation presented in \autoref{sec:mathematics}, an
assumption is made that the number of bets is very high (tends to infinity).
This work does not try to justify this assumption in a strong mathematical
form~\footnote{It is on the author's TODO list to try and prove this.}. Here's a
soft, non-mathematical reasoning on why the author thinks this is fine: The
assumption is made in order to write the equations in terms of probabilities
instead of the number of outcomes divided by the number of bets. Therefore, as
long as the input probabilities are \textit{conservatively} estimated, the
framework should still be valid. This essentially represents the most important
margin of safety~\cite{intelligentInvestor}.

@@

\noindent where $\mathcal{A}$ is the value of assets (capital), $N_c$ is the
number of candidate companies to consider, $f_j$ is the allocated fraction to
$j$th company, and $k_j$ is a return for a company $j$ defined as the relative
difference between the estimated intrinsic value under a given scenario
($\mathcal{V}$) and the market capitalization at the time of investing
($\mathcal{M}$):
\begin{equation}
\label{eq:2}
    k_j = \frac{\mathcal{V}_j - \mathcal{M}_j}{\mathcal{M}_j}
\end{equation}

@@

\noindent where $n_i \in {n_{i1}, n_{i2}, \hdots, n_{N_o}}$ is the number of
times the $i$th outcome has occurred. Note that $k_{ij}$ represents the return
of the $j$th company for the $i$th outcome. Following original Kelly's approach,
a logarithmic growth function $\mathcal{G}$ is introduced:
\begin{equation}
\label{eq:4}
    \mathcal{G} = \lim_{N_a \to \infty} \frac{1}{N_a} \ln
        \frac{\mathcal{A}_{N_a}}{\mathcal{A}_0}
\end{equation}

@@

\noindent If one assumes an infinite number of (re)allocations
$N_a$\footnote{The assumption regarding the infinite number of allocations is
something that the author is slightly uncomfortable with, but \textit{feels} it
is fine because of the margins of safety embedded into the thinking that goes
into assessing each investment opportunity.}, the following relation holds:
\begin{equation}
\label{eq:7}
    \lim_{N_a \to \infty} \frac{n_i}{N} = p_i
\end{equation}

@@

\begin{table}[!ht]
\caption{Company E with current market cap of 17.6M CAD.}
\vspace{0.25cm}
\centering
\begin{tabular}{l|c|c}
Scenario & Intrinsic value & Probability \\
\hline
Total loss & 0 CAD & 5\% \\
Bear thesis & 10M CAD & 25\% \\
Base thesis & 25M CAD & 70\% \\
\end{tabular}%
\label{tab:companyE}%
\end{table}%

\begin{table}[!ht]
\caption{Company F with current market cap of 581M USD.}
\vspace{0.25cm}
\centering
\begin{tabular}{l|c|c}
Scenario & Intrinsic value & Probability \\
\hline
Total loss & 0 USD & 20\% \\
Base thesis & 1B USD & 80\% \\
\end{tabular}%
\label{tab:companyF}%
\end{table}%

@@

\noindent The last item is particularly interesting to the author. According to
Actuarial Life Tables in~\cite{lifeTables}, the probability of the (currently 33
year--old) author dying within the next year is approximately 0.25\%. That is
two orders of magnitude higher than the probability of the permanent loss of
capital for this portfolio. Considering that the portfolio has six stocks, that
is a very strong argument against excessive diversification, especially if:
\begin{itemize}
    \item One thinks of stocks as ownership shares of businesses, which implies
    long--term thinking and not being bothered by market fluctuations, 
    \item One embeds a margin of safety in different scenarios for different
    companies by e.g. recognizing that both unknown and known bad things may
    happen.
\end{itemize}

\noindent The observation about excessive diversification is inline with the
thoughts from the Poor Charlie's Almanack~\cite{almanack}.
